---
title: Comparing the "lm" function from *stats* package with "train" function from *caret*
  package
author: "Bishnu Poudel"
date: "May 23, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning= FALSE, message= FALSE)
```

## Executive Summary

In this presentation, I will compare the *outputs and coefficients* of the **lm** function from the **stats** package to the **train** function from the **caret** package. 

I will run these two functions on the same data in order to generate two similar linear models. 

- Here, in these slides we will compare the coefficients and p values of the resulting model.

- Similarly, in the shiny web application I built, we can manually enter the input values, in order to compare the outputs/predictions of these models.


Codes for the assigment are in the github repository below:

## Coefficients of the models
```{r stats, echo = TRUE,include=FALSE}
mtcarsFac<-mtcars[,colnames(mtcars) %in% c("mpg","wt","qsec","am")]
  mtcarsFac$am<- as.factor(mtcarsFac$am)
library(caret);library(shiny); library(ggplot2)
```
```{r, echo=TRUE}
summary( lm(mpg~wt+qsec+am, data=mtcarsFac))$coef  # stats package
summary(train(mpg~wt+qsec+am, data=mtcarsFac, method="glm"))$coef #caret
```


## Including Server calculations
Since I was not able to embed the application itself into the presentation, explaining the server calcuations here.
*We will need shiny, ggplot2 and caret packages*
```{r, echo = TRUE, eval=FALSE}
library(shiny);  library(ggplot2);library(caret)
```
 *We create and store the two models. Model prediction is done inside a reactive clause, repeated for the modGLM model as well. The outputs of the model goes to the user interface.*
```{r, echo = TRUE, eval=FALSE}
  modLM<-lm(mpg~wt+qsec+am, data=mtcarsFac)
  modGLM<-train(mpg~wt+qsec+am, data=mtcarsFac, method="glm")
  model1pred <- reactive({
    predict(modLM, newdata= data.frame(wt=weight,qsec=quarterTime
                                       ,am=automatic_manual))})
```



## Conclusion

As we witnessed the coefficients and p values are exactly same. 
Therefore the two approaches are identical.

Now, we can go ahead and test the results by providing actual inputs in our shiny web application!



